---
title: "master"
---

## Set the directory 
```{r dir, echo=TRUE}
setwd("F:/Habeba/Data Science/Slides/5.Reproducible Research/Week 2/Assignment")
```

## Read the file 
```{r read_dir, echo=TRUE}
activity_dataset <- read.csv("activity.csv")
```
## Claculate total steps for each day
```{r tot_stps, echo=TRUE}
total_steps <- aggregate(activity_dataset$steps , list(datee = activity_dataset$date), sum)
```

## Creating Hist plot 
```{r hist_plot, echo=TRUE}
hist(total_steps$x , main= "Total steps per day")
```


## Calculating mean and median for total steps per day
```{r mean_med, echo=TRUE}
total_steps_mean <- mean(total_steps$x , na.rm = TRUE)
total_steps_median <- median(total_steps$x , na.rm = TRUE)
```
The mean is `r total_steps_mean` and the median is `r total_steps_median`.

## For creating Time Series Plot
```{r time_series_plot, echo=TRUE}
library(ggplot2)
avf_steps <- aggregate(steps~interval,activity_dataset, mean)
ggplot(data = avf_steps , aes(x=interval, y = steps) , type="I", main= "Average Daily Activity Pattern" , xlab="5-minute Interval", ylab= "Average Number of Steps") + geom_line()
```

## Calculate maximum of average steps per day
```{r maxi_steps, echo=TRUE}
 max_mean_steps <- avf_steps[which.max(avf_steps$steps),]
```
The mmaximum steps is `r max_mean_steps`

##For calculating total NA and remove them
```{r NA, echo=TRUE}
 missingVals <- is.na(activity_dataset$steps)
withoutNA_activity_dataset <- transform(activity_dataset,steps = ifelse(is.na(activity_dataset$steps),avf_steps$steps[match(activity_dataset$interval,avf_steps$interval)],activity_dataset$steps))
total_steps_withoutNA <- aggregate(steps~ date , withoutNA_activity_dataset, sum)
hist(total_steps_withoutNA$steps , main= "Total steps per day without null")
```


## Calculate mean and median of data without NA
```{r NA_mean_med, echo=TRUE}
 total_steps_mean_withoutNA <- mean(total_steps_withoutNA$steps , na.rm = TRUE)
total_steps_median_withoutNA <- median(total_steps_withoutNA$steps , na.rm = TRUE)
```
The mean is `r total_steps_mean_withoutNA` and the median is `r total_steps_median_withoutNA`.

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r comparing, echo=TRUE}
DayType <- function(date)
 {
   day <- weekdays(date)
   if (day %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'))
       return ("weekeday")
   else if (day %in% c('Saturday', 'Sunday'))
return ("weekend")
   else
       stop ("Invalid Date Format.")
}
activity_dataset$date <- as.Date(activity_dataset$date)
activity_dataset$day <- sapply(activity_dataset$date, DayType)
meanStepsByDay <- aggregate(steps ~ interval + day, activity_dataset, mean)
ggplot(data = meanStepsByDay, aes(x = interval, y = steps) , main = "Average Daily Activity Pattern" , xlab = "5-minute Interval" , ylab = "Average Number of Steps" ) +
  geom_line() +
  facet_grid(day ~ .) 
```
